Characterizing subarctic biomes for land surface modeling of pollution and climate risk

### 2021-03-04 VegOzone.tex
First draft of results.

The density distribution for Svanvik and the respective \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, and \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}} (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) display a similar pattern as the northern Fennoscandic climatology. The spring peak apparently occurs slightly earlier (in the beginning of April). As Svanvik is located at lower altitude than the other sites, the earlier decline of the ozone peak may point to a stronger ozone uptake by vegetation. \chem{CO_2} uptake coniferous trees can occur as early as doy 100 \citep{TB:Kolari2007}. The average ozone concentration at Svanvik is lower than at the other sites, as already mentioned, which is in league with a decrease in tropospheric ozone with decreasing altitude. In July--September, ozone is occasionally almost completely depleted during night time.

@@

\subsection{Model parameters}
From Fig.~\ref{fig:svanhovd_research_station}, we roughly estimated land use types for an approximately $1\times 1\,\unit{km}$ patch centered at the NIBIO Environment Centre Svanhovd : $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban.

Default PFTs are derived from \citet{ICP:MappingManual2017} and will be referred to as \emph{mapping manual} abbreviated as MM. A comprehensive list of these model parameters can be found in Supplement~Table~S1.
Initial test simulations with the mapping manual derived PFTs showed unrealistically low $g_\mathrm{sto}$ in 2019 particularly for grassland (Appendix~Fig.~\ref{fig:pody_mm_composit}). Only during an extended period in late July did we find substantial stomatal conductance. In ecological terms, this would mean that there was almost no growth of grass in the summer of 2019 -- a prediction which is easily falsified by reality. We identified $f_\mathrm{T}$ as abnormally low, being the limiting factor of stomatal conductance. For these reasons, we propose bespoke parameterizations for the main vegetation types Norway spruce, birch, and perennial grassland which are better adapted to the climatic conditions in the target region.

Note that these parameterizations represent hypothetical PFTs or species that have yet to be verified by experiments.

We hypothesize that perennial vegetation will likely prioritize their growth or fast adolescence throughout the short, subarctic growing season and hence their carbon sequestration. Since the $\mathrm{DO_3SE}$ in its current state does not simulate $A_\mathrm{n}$, we assume for our purpose a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011}. We then tune the species temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) separately for higher stomatal conductance. We illustrate this procedure graphically in Fig.~\ref{fig:f_temp_grassland} for grassland. We aim to increase the enclosed area between probability density function (PDF) and the respective response function. For comparison, we define two climatological references 1990s and 2000s. For Norway spruce and birch see Appendix Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}.

We assume two different types of hypothetical temperature acclimations: \emph{Boreal} and \emph{cold}. For either, we further assume that the hypothetical species are more tolerant to cold conditions than MM suggests and shift $f_\mathrm{T}$ towards colder temperatures. We construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to MM. In the same way, \emph{boreal} refers to a hypothetical species which is very tolerant to cold but is sensitive to high temperatures, and most efficient at cool temperatures ($T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000). For perennial grassland these are depicted together with the temperature PDF in Fig.~\ref{fig:f_temp_grassland}a).

Norway spruce was experimentally found to be active already at rather low air temperatures and could reach $60\,\unit{\%}$ photosynthetic activity as early as doy~100 \citet{TB:Wallin2013, TB:Kolari2007}. We derived temperature acclimation, based on time series of observed \chem{CO_2} uptake and temperature at two observation sites in Southern ($\rightarrow$\emph{cold}) and Northern ($\rightarrow$\emph{boreal}) Finland \citet{TB:Kolari2007}. We assume a similar acclimation for birch. All bespoke temperature parameters are comprehensibly listed in Table~\ref{tab:sensitivity_tests_temp}.\\

For an acclimation with respect to light conditions, we hypothize that the light threshold for a $50\,\unit{\%}$ opening of plants' stomata may differ in subarctic species. We analytically derive the inverse function $f_{\mathrm{light},k}^{-1}$ for each PFT or species $k$
\begin{equation}
  \gamma_k(f_\mathrm{light}) := f_{\mathrm{light}, k}^{-1} = -\frac{\ln(1-f_\mathrm{light})}{\alpha_k}.
  \label{eq:inverse_function}
\end{equation}
We solve Eq.~(\ref{eq:inverse_function}) for the mapping manual default value for each PFT or species $\alpha_\mathrm{MM}$ at its halfway value $0.5$ 
\begin{equation}
  \gamma(0.5) = -\frac{\ln(0.5)}{\alpha_\mathrm{MM}},
  \label{eq:inverse_function_halfway}
\end{equation}
and define a variation $\gamma' = \gamma \cdot \epsilon$ with $\epsilon \in \{0.8, 1.2\}$. We resolve Eq.~(\ref{eq:inverse_function}) again for $\alpha$ and find $\alpha'(\gamma')$. The resulting functions are shown in Fig.~\ref{fig:f_temp_grassland}b). All derived parameters are tabulated in Table~\ref{tab:sensitivity_tests_light}.

@@

\begin{table}[t]
  \caption{Bespoke temperature parameterizations. MM refers to mapping manual \citep{GCB:Mills2010,ICP:MappingManual2017}.}
  \label{tab:sensitivity_tests_temp}
  \begin{tabular}{llrrr}
    \tophline
    Species & type & $T_\mathrm{min}$ & $T_\mathrm{opt}$ & $T_\mathrm{max}$ \\
    \middlehline
    \multirow{3}{*}{Spruce} & MM & 0 & 20 & \multirow{3}{*}{100}\\
    & Boreal & 0 & 10 & \\
    & Cold & 0 & 15 & \\
    \middlehline
    \multirow{3}{*}{Birch} & MM & 5 & 20 & \multirow{3}{*}{100}\\
    & Boreal & 0 & 10\\
    & Cold & 5 & 15\\
    \middlehline
    \multirow{3}{*}{Grassland} & MM & 10 & 24 & 36\\
    & Boreal & 0 & 10 & 24 \\
    & Cold & 5 & 15 & 36\\
    \bottomhline
    \end{tabular}
\end{table}

\begin{table}[t]
  \caption{Bespoke light parameterizations. MM refers to mapping manual \citep{GCB:Mills2010,ICP:MappingManual2017}.}
  \label{tab:sensitivity_tests_light}
  \begin{tabular}{llr}
    \tophline
    Species & type & $\alpha$\\
    \middlehline
    \multirow{3}{*}{Spruce} & MM & 0.006 \\
    & PPFD0.8 & 0.008  \\
    & PPFD1.2 & 0.005  \\
    \middlehline
    \multirow{3}{*}{Birch} & MM & 0.004 \\
    & PPFD0.8 & 0.005\\
    & PPFD1.2 & 0.004\\
    \middlehline
    \multirow{3}{*}{Grassland} & MM & 0.011\\
    & PPFD0.8 & 0.014 \\
    & PPFD1.2 & 0.009 \\
    \bottomhline
    \end{tabular}
\end{table}

To quantify the resulting bespoke parameterizations, we propose the following metric.
{\bf TODO: Write about Boreal and cold parameter sets and their justification. Write about suggested method for optimizing PFTs in a simple but physiological way based on optimizing maximum climatological mean $g_\mathrm{sto}$ and morning responds. If plants have a short growing season up north optimizing \chem{CO_2} uptake around noon (highest light intensity) and speed up opening of stomata at lower intensities (morning), additional contained lower climatological variance? An idea to justify our choice based on the full set of climatological data for Svanvik is shown in Fig.~\ref{fig:javis_func_opt_mean}.}

\begin{figure}[t]
  %\includegraphics[width=8.3cm]{javis_func_opt_mean}
  \includegraphics[width=8.3cm]{fig10}
  \caption{Proposed metric to test bespoke response functions. GS (May--August) climatological averages and standart deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time) are shown for (a) birch; (b) Norway spruce; (c) perennial grassland. For simplicity, $f_\mathrm{SW}$ is neglected in this calculation. }
  \label{fig:javis_func_opt_mean}
\end{figure}

Start ($A_\text{start}$) and end ($A_\text{end}$) of growing season for each PFT are given in units of day of year (doy) and estimated as follows. Evergreen needle leaf trees show photosynthetic activity already at low temperatures \citep{TP:Wallin2013, TB:Kolari2007}. We use a MODIS (Aqua/Terra) retrieved net photosynthesis product for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine onset and end of growing season. Net photosynthesis data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}). By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{doy}$ for 2018 and day 106 in 2019. $A_\text{end}$ amounts to day 261 and 274, respectively, and shall be used for all vegetation alike. The resulting growing season for Norway spruce in 2019 was one month longer then in 2018. For estimating $A_\text{start}$ of birch tree, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Fig.~\ref{fig:greening_season_change_Svanvik}). We find 129 and 130, respectively. From the latitude model typically used for localizing $\mathrm{DO_3SE}$, we get $A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$ indicating an about $2\,\unit{months}$ longer growing season. For perennial grassland, we assume a latency period of about $1\,\unit{month}$ after snow melt. Data for first snow-free day is taken from the weather station at Øvre Neiden in Sør-Varanger. In both years, the first snow-free day incidents with the estimated date of bud burst. As leafs of birches in northern Fennoscandia \emph{Betula pubescens} are smaller than southern specimen, we have measured their dimension for the purpose of this study and find $(3.0\pm 0.5)\,\unit{cm}$. \citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area.

As pointed out by \citet{ACP:Bueker2012}, a very important factor in deriving a proper $\mathrm{POD_y}$ is the soil texture or how much water the soil can hold. As Svanhovd is located in the bed of the Pasvik river, we assume a sandy loam texture at $60\,\unit{cm}$ depth.\\

\subsection{POD and ozone impact}

We have modeled $\mathrm{POD_y}$ for three natural/semi-natural vegetation PFTs, birch, Norway spruce, and perennial grassland. The results are shown comprehensively in Fig.~\ref{fig:pody} as a function of doy for March--October. A flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) has been chosen. \chem{[O_3]} are plotted on the same axis and scales as $G_\text{sto}^\text{leaf}$ but in units of $\unit{ppb}$.

As expected, $\mathrm{POD_y}$ is higher in 2018 than in 2019 for all PFTs. In the end of the 2018 season, Birch took up $13.1\,\unit{mmol\,m^{-2}\,PLA}$, while perennial grassland had an uptake of $4.5\,\unit{mmol\,m^{-2}\,PLA}$. The largest interannual variation is also found in Birch ($2.1\,\unit{mmol\,m^{-2}\,PLA}$) and perennial grassland ($2.6\,\unit{mmol\,m^{-2}\,PLA}$). Whereas, there is only a small difference for Norway spruce ($0.3\,\unit{mmol\,m^{-2}\,PLA}$) between the two years.

%Except for perennial grassland, model integrations with climatological forcings show higher $\mathrm{POD_y}$ than 2018. This can be explained by means of non linearity in modeled stomatal conductance (Eq.~(\ref{eq:stomatal})) in Eq.~(\ref{eq:pod}). The climatology in $\mathrm{POD_y}$ therefore represents rather an ideal year for growth then a true climatology. To establish a true climatology, it would be necessary to run the $\mathrm{DO_3SE}$ model for 10--20 years with respective observed phenology as well as meteorological and ozone forcing.

%Based on this interpretation of climatology in $\mathrm{POD_y}$, we can learn more about potential ozone damage. The time derivative $\frac{\mathrm{\Delta POD_y}}{\Delta t}$ displays very different features for each PFT and meteorological forcing. Although Norway spruce shows a low interannual variation, $\frac{\mathrm{\Delta POD_y}}{\Delta t}$ (Fig.~\ref{fig:pody_rel}) indicates an extended period of higher than average ozone uptake in spring 2018 (steep slope), while ozone uptake in June/July was anomalously low in 2019 (shallow slope) but high in fall. This is most prominently visible in Fig.~\ref{fig:pody}h) for perennial grassland, where consecutive periods of low $G_\text{sto}^\text{leaf}$ manifest in three plateaus in $\mathrm{POD_y}$ throughout the season (1: $184-198\,\unit{doy}$, 2: $209-228\,\unit{doy}$, 3: $235-240\,\unit{doy}$). An episode with anomalous high ozone in mid June ($194\,\unit{doy}$) can be identified but does not amount to unusually high ozone uptake on an annual basis.

@@

\begin{table}[t]
  \caption{Estimated total biomass reduction in \unit{\%} due to \chem{O_3} following \citet{ICP:MappingManual2017}.}
  \label{tab:biomass_reduction}
\begin{tabular}{lccrr}
\tophline
Year & \multicolumn{4}{c}{PFT}\\
& Birch & Norway spruce & \multicolumn{2}{c}{Perennial grassland}\\
\middlehline
2018 & 11.99 & 2.34 & 8.06 & 10.51$^*$\\
2019 & 10.07 & 2.26 & 6.43 & 7.90$^*$\\
clim. & 13.81 & 2.58 & 6.67 & 8.29$^*$\\
\bottomhline
\end{tabular}
\belowtable{$^*$ above ground biomass} % Table Footnotes
\end{table}

\begin{figure}[t]
  %\includegraphics[width=8.3cm]{do3se_results_pod_v3_Birch}\\
  %\includegraphics[width=8.3cm]{do3se_results_pod_v3_Spruce}\\
  %\includegraphics[width=8.3cm]{do3se_results_pod_v3_Grassland}
  \includegraphics[width=8.3cm]{fig11}
  \caption{Comprehensive comparison of $\mathrm{DO_3SE}$ modeling results with default parameterization from mapping manual (MM)and bespoke parameters for Svanhovd. $\mathrm{POD_1}$ is shown for the different temperature response functions (MM, boreal, cold) and light thresholds (MM, PPFD0.8, PPFD1.2) at the end of a MM GS (lighter hue and dotted lines) and a derived bespoke GS. (a) Birch; (b) Norway spruce; (c) Perennial grassland.}
  \label{fig:pody_rel}
\end{figure}

@@

\begin{figure*}[t]
  \includegraphics[width=12cm]{figB1}
  \caption{Estimated begin (BGS) and end (EGS) of growing season for coniferous trees from MODIS Aqua/Terra net photosynthesis (PSN) product. A $1\times 1\,\unit{km}$ area around Svanhovd was selected. Daily averaged data for both 2018 and 2019 has been fitted with a quadratic polynomial function. The numerically computed root yields: BGS 122/106 day of year (doy) and EGS 261/274 day of year for 2018/2019, respectively.}
  \label{fig:modis_Psn}
\end{figure*}

>>>>

The density distribution for Svanvik and the respective \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, and \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}} (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) display a similar pattern as the northern Fennoscandic climatology. The spring peak apparently occurs slightly earlier (in the beginning of April). As Svanvik is located at lower altitude than the other sites, the earlier decline of the ozone peak may point to a stronger ozone uptake by vegetation. \chem{CO_2} uptake coniferous trees can occur as early as day of the year (\unit{doy}) 100 \citep{TB:Kolari2007, TP:Wallin2013}. The average ozone concentration at Svanvik is lower than at the other sites, as already mentioned, which is in league with a decrease in tropospheric ozone with decreasing altitude. In July--September, ozone is occasionally almost completely depleted during night time.

@@

\subsection{Model parameters}
\label{subsec:do3se_parameters}

From Fig.~\ref{fig:svanhovd_research_station}, we roughly estimated land use types for an approximately $1\times 1\,\unit{km}$ patch centered at the NIBIO Environment Centre Svanhovd : $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban.

Default parameters for these species or PFTs are derived from \citet{ICP:MappingManual2017} and will be referred to as \emph{mapping manual} abbreviated as MM in the following. A comprehensive list of model parameters can be found in Supplement~Table~S1.
Initial test simulations with the MM PFTs showed unrealistically low $g_\mathrm{sto}$ in 2019 particularly for perennial grassland (Appendix~Fig.~\ref{fig:pody_mm_composit}). Only during an extended warm period in late July did we find substantial stomatal conductance. In ecological terms this would mean that there was almost no growth of grass in the summer of 2019 -- a prediction which is easily falsified by reality. We identified $f_\mathrm{T}$ as abnormally low, being the limiting factor of stomatal conductance. Therefor, we propose bespoke parameterizations which may describe an actual acclimation of these species to the climatic conditions in the target region.

Note, however, that these parameterizations represent hypothetical PFTs or species that have yet to be verified by experiments.

We hypothesize that perennial vegetation will likely prioritize their growth or fast adolescence during a short, subarctic growing season and hence their carbon sequestration. Because the $\mathrm{DO_3SE}$ model currently does not simulate $A_\mathrm{n}$, we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$. The procedure is graphically illustrated in Fig.~\ref{fig:f_temp_grassland} for grassland. We aim to increase the enclosed area between probability density function (PDF) and the respective response function. We show two climatological reference periods 1990s and 2000s to indicate that the vegation is subject to ongoing climate change and acclimations may be changing. Resulting reponse functions for Norway spruce and downy birch are displayed in Appendix~Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}.

We define two different hypothetical temperature acclimations: \emph{Boreal} and \emph{cold}. For either, we assume that the hypothetical species are more tolerant to cold conditions and shift $f_\mathrm{T}$ towards colder temperatures accordingly. We construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to MM. In the same way, \emph{boreal} refers to a hypothetical species which is very tolerant to cold but is sensitive to high temperatures, and most efficient at cool temperatures ($T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000). For perennial grassland the resulting response functions are depicted together with the PDF of temperatur in Fig.~\ref{fig:f_temp_grassland}a).

Norway spruce was experimentally found to be active already at rather low air temperatures and could reach $60\,\unit{\%}$ photosynthetic activity as early as \unit{doy}~100 \citep{TB:Kolari2007, TP:Wallin2013}. Based on the time series of observed \chem{CO_2} uptake and temperature at two observation sites in Southern ($\rightarrow$ \emph{cold}) and Northern ($\rightarrow$ \emph{boreal}) Finland \citep{TB:Kolari2007}, we estimate $10\,\unit{^\circ C}\le T_\mathrm{opt} \le 15\,\unit{^\circ C}$. We assume a similar acclimation for downy birch. All bespoke temperature parameters are listed in Table~\ref{tab:sensitivity_tests_temp}.\\

For an acclimation with respect to light conditions, we hypothize that the light threshold for a $50\,\unit{\%}$ opening of plants' stomata may differ in subarctic species. We analytically derive the inverse function $f_{\mathrm{light},k}^{-1}$ for each PFT or species $k$
\begin{equation}
  \gamma_k(f_\mathrm{light}) := f_{\mathrm{light}, k}^{-1} = -\frac{\ln(1-f_\mathrm{light})}{\alpha_k}.
  \label{eq:inverse_function}
\end{equation}
First, we solve Eq.~(\ref{eq:inverse_function}) for the MM default value $\alpha_\mathrm{MM}^k$ for each PFT or species $k$ at $0.5$ 
\begin{equation}
  \gamma^k(0.5) = -\frac{\ln(0.5)}{\alpha_\mathrm{MM}^k},
  \label{eq:inverse_function_halfway}
\end{equation}
and define a variation $\gamma' = \gamma \cdot \eta$ with $\eta \in \{0.8, 1.2\}$. We invert Eq.~(\ref{eq:inverse_function}) for $\alpha$ and find $\alpha'(\gamma')$. The resulting functions for perennial grassland are shown in Fig.~\ref{fig:f_temp_grassland}b) (for downy birch and Norway spruce refer Appendix~Figs.~\ref{fig:f_temp_spruce}b)--\ref{fig:f_temp_birch}b)). All derived parameters are tabulated in Table~\ref{tab:sensitivity_tests_light}.

@@

\begin{table}[t]
  \caption{Bespoke temperature parameterizations. MM refers to mapping manual \citep{GCB:Mills2010,ICP:MappingManual2017}.}
  \label{tab:sensitivity_tests_temp}
  \begin{tabular}{llrrr}
    \tophline
    Species & type & $T_\mathrm{min}$ & $T_\mathrm{opt}$ & $T_\mathrm{max}$ \\
    \middlehline
    \multirow{3}{*}{Downy birch} & MM & 5 & 20 & \multirow{3}{*}{100}\\
    & Boreal & 0 & 10\\
    & Cold & 5 & 15\\
    \middlehline
    \multirow{3}{*}{Norway spruce} & MM & 0 & 20 & \multirow{3}{*}{100}\\
    & Boreal & 0 & 10 & \\
    & Cold & 0 & 15 & \\
    \middlehline
    \multirow{3}{*}{Perennial grassland} & MM & 10 & 24 & 36\\
    & Boreal & 0 & 10 & 24 \\
    & Cold & 5 & 15 & 36\\
    \bottomhline
    \end{tabular}
\end{table}

\begin{table}[t]
  \caption{Bespoke light parameterizations. MM refers to mapping manual \citep{GCB:Mills2010,ICP:MappingManual2017}.}
  \label{tab:sensitivity_tests_light}
  \begin{tabular}{llr}
    \tophline
    Species & type & $\alpha$\\
    \middlehline
    \multirow{3}{*}{Downy birch} & MM & 0.004 \\
    & PPFD0.8 & 0.005\\
    & PPFD1.2 & 0.004\\
    \middlehline
    \multirow{3}{*}{Norway spruce} & MM & 0.006 \\
    & PPFD0.8 & 0.008  \\
    & PPFD1.2 & 0.005  \\
    \middlehline
    \multirow{3}{*}{Perennial grassland} & MM & 0.011\\
    & PPFD0.8 & 0.014 \\
    & PPFD1.2 & 0.009 \\
    \bottomhline
    \end{tabular}
\end{table}

As indicated above, a better acclimation will result in a higher average $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$. To quantify the resulting bespoke parameterizations, we propose the following metric. For a generic GS (May--August), we compute climatological average and standart deviation of $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal})) at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time). We presume that relative stomatal conductance of around noon (highest light intensity) and in the early morning (lower intensities) are good proxies for \chem{CO_2} uptake efficiency. Our arbitrary optimization target is $60\,\unit{\%}$. Additional constrains are a small deviation between noon and morning and a low standard deviation indicating a higher robustness to interannual variability of growing conditions. For simplicity, we neglect the dependency on SWP in this regard. From the results shown in Fig.~\ref{fig:javis_func_opt_mean}, we identify boreal, PPFD0.8 as the best acclimation for all hypothetical species or PFTs alike. Overall, Norway spruce displays the least deviation between the different types of parameterizations, while the divergance for perannial grassland is substantial.

\begin{figure}[t]
  %\includegraphics[width=8.3cm]{javis_func_opt_mean}
  \includegraphics[width=8.3cm]{fig10}
  \caption{Proposed metric to test bespoke response functions. GS (May--August) climatological averages and standart deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time) are shown for (a) Downy birch; (b) Norway spruce; (c) perennial grassland. For simplicity, $f_\mathrm{SW}$ is neglected in this calculation. }
  \label{fig:javis_func_opt_mean}
\end{figure}

Another important factor for ozone exposure metrices is the timing and length of the GS. Start ($A_\text{start}$) and end ($A_\text{end}$) of GS for each PFT are given in units of \unit{doy} (Tab.~\ref{tab:sensitivity_tests_gs}) and estimated as follows. Evergreen needle leaf trees show photosynthetic activity already at low temperatures \citep{TP:Wallin2013, TB:Kolari2007}. We use a MODIS (Aqua/Terra) retrieved net photosynthesis product for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine $A_\text{start}$ and $A_\text{end}$. Net photosynthesis data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}). By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{doy}$ for 2018 and \unit{doy} 106 in 2019. $A_\text{end}$ amounts to \unit{doy} 261 and 274, respectively, and shall be used for all vegetation alike. The resulting growing season for Norway spruce in 2019 was one month longer then in 2018. For estimating $A_\text{start}$ of downy birch, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Fig.~\ref{fig:greening_season_change_Svanvik}). We find 129 and 130, respectively. For both years, these dates coincident with the first snow-free day at the meterological weather station at Øvre Neiden (Sør-Varanger, NOR) which is relatively close-by. From the latitude model typically used for localizing $\mathrm{DO_3SE}$, we get $A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$ indicating an about $2\,\unit{months}$ longer growing season. Due to a lack of quantitative field observation, we assume a latency period of about $1\,\unit{month}$ after snow-melt for perennial grassland.

\begin{table}[t]
  \caption{Bespoke start and end of growing season in units of \unit{doy}. MM refers to mapping manual \citep{GCB:Mills2010,ICP:MappingManual2017}.}
  \label{tab:sensitivity_tests_gs}
  \begin{tabular}{llcc}
    \tophline
    Species & Year & $A_\mathrm{start}$ & $A_\mathrm{end}$\\
    \middlehline
    \multirow{2}{*}{Downy birch} & 2018 & 129$^*$ & 261$^\dagger$ \\
    & 2019 & 130$^*$ & 274$^\dagger$ \\
    \multirow{2}{*}{Norway spruce} & 2018 & 122$^\dagger$ & 261$^\dagger$ \\
    & 2019 & 106$^\dagger$ & 274$^\dagger$ \\
    \multirow{2}{*}{Perennial grassland} & 2018 & 159$^\ddagger$ & 261$^\dagger$\\
    & 2019 & 161$^\ddagger$ & 274$^\dagger$ \\
    \bottomhline
  \end{tabular}
  \belowtable{$^\dagger$ MODIS (Aqua/Terra) net photosynthesis product;\\
    $^*$ $5\,\unit{days}-5\,\unit{^\circ C}$-rule, $T_\mathrm{air}$ from \href{seNorge.no}{seNorge.no};\\
    $^\ddagger$ One month after snow melt; reference station Øvre Neiden (Sør-Varanger, NOR)} % Table Footnotes
\end{table}

As pointed out by \citet{ACP:Bueker2012}, a very important factor in deriving a proper $\mathrm{POD_y}$ is the soil texture or how much water the soil can hold. As Svanhovd is located in the bed of the Pasvik river, we assume a sandy loam texture at $60\,\unit{cm}$ depth. Further, leaves of downy birch in northern Fennoscandia are smaller than southern specimens'. Here, we have measured their dimension and found $(3.0\pm 0.5)\,\unit{cm}$. \citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area.

\subsection{$\mathrm{POD_1}$ results and implications on ozone impact}
\label{subsec:do3se_results}

We have modeled $\mathrm{POD_y}$ (flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA)) for three natural/semi-natural vegetation types, downy birch, Norway spruce, and perennial grassland using both MM parameterizations and a variaty of bespoke parameterizations in $f_\mathrm{T}$, $f_\mathrm{light}$, and GS (described in detail in Section~\ref{subsec:do3se_parameters}). The results are shown comprehensively in Fig.~\ref{fig:pody_rel}. For comparison, critical levels (CLs) for a $4\,\unit{\%}$ ($2\,\unit{\%}$ for Norway spruce) reduction in biomass in accordance to \citet{ICP:MappingManual2017} are shown as horizontal dashed lines. Different years are color coded. Lighter hue indicates simulations with MM GS and MM canopy and leaf parameters. Annotated $\pm$ refers to bespoke light response functions, while bespoke temperature response functions are indicated by categories on the x-axis. Closed/open symbols represent SWP sensitivity switched off/on in the model.

For downy birch (Fig.~\ref{fig:pody_rel}a), all simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and display the largest spread among the species ($16\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 for boreal type). As expected, SWP reduces ozone uptake in 2018 mainly. The uncertainty induced by the light sensitivity threshold ranges between $(1.5..2.5)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $(-2.2..-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. With MM GS, 2018 displays increasingly larger ozone uptake than 2019 with more acclimated paramters. However accounting for the shorter and later GS in 2018, we find a reversal of this relationship in the boreal type. We deduce an overall systematic uncertainty of $xx$.

For Norway spruce (Fig.~\ref{fig:pody_rel}b), most simulations exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). Norway spruce displays an overall lower spread ($10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 for boreal type) than downy birch, but appears more sensitive to GS and interannual veriability in conditions (boreal GS MM compared to GS bespoke). The sensitivity to the light threshold is also smaller compared to downy birch. Accounting for the shorter and later GS in 2018, we find again a reversal between predicted ozone damage in 2018 and 2019 for the boreal type. We deduce an overall systematic uncertainty of $xx$.

For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations with bespoke GS stay below the CL. CL is only exceeded for boreal types and cold type in 2018. We find the same reveral between 2018 and 2019 as for the other species in the boreal type. Perannial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant for the MM GS. The spread in 2018 for boreal type is of the order of $10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. We deduce an overall systematic uncertainty of $xx$.

These results allow for an assessment of a large variety of systematic uncertainties in the exposure-based risk metric for ozone induced damage in subarctic vegetation. In summary, we identify the following main effects on the model predicted ozone uptake:
\begin{itemize}
\item Optimal acclimation of species or PFT $\rightarrow$ increased uptake in any year,
\item Identification of proper timing of the growing season $\rightarrow$ reduced uptake in any year,
\item Drought conditions $\rightarrow$ reduced uptake,
\item Early onset of growing season $\rightarrow$ increased uptake.
\end{itemize}
The exact order of magnitude of these effects varies with species or PFT and from year to year. The systematic uncertainty induced by our proposed temperature acclimation is, in general, larger than for a $\pm 20\,\unit{\%}$ variation in light response threshold and can likely exceed $xx\,\unit{\%}$ for some years and species. This would lead to a huge underestimation of induced ozone damage. At the same time, the magnitude of light response uncertainty is increasing with acclimation to colder temperature. The impact of the timing of the GS can be at the same oder of magnitude as the acclimation of response function but points into the oposite direction. Especially, the timing of $A_\mathrm{start}$ with respect to the ozone spring peak (compare 2018 GS MM and GS bespoke) induces large uncertainties which are increasing if an acclimation to colder climate is anticipated.

\begin{figure}[t]
  \includegraphics[width=8.3cm]{fig11}
  \caption{$\mathrm{DO_3SE}$ modeling results for $\mathrm{POD_1}$ with default parameterization (MM) and bespoke parameters for Svanhovd. Bespoke parameterizations include temperature response functions for boreal, cold and light thresholds (PPFD0.8, PPFD1.2) at the end of a MM GS (lighter hue and dotted lines) and the estimated GS for Svanhoved (2018, 2019). (a) Downy birch; (b) Norway spruce; (c) Perennial grassland. The dashed horizontal lines indicate the critical levels defined in \citet{ICP:MappingManual2017} for each species, respectively.}
  \label{fig:pody_rel}
\end{figure}

@@

\begin{table}[t]
  \caption{Estimated total biomass reduction in \unit{\%} due to \chem{O_3} following \citet{ICP:MappingManual2017}.}
  \label{tab:biomass_reduction}
\begin{tabular}{lccrr}
\tophline
Year & \multicolumn{4}{c}{PFT}\\
& Downy birch & Norway spruce & \multicolumn{2}{c}{Perennial grassland}\\
\middlehline
2018 & 11.99 & 2.34 & 8.06 & 10.51$^*$\\
2019 & 10.07 & 2.26 & 6.43 & 7.90$^*$\\
clim. & 13.81 & 2.58 & 6.67 & 8.29$^*$\\
\bottomhline
\end{tabular}
\belowtable{$^*$ above ground biomass} % Table Footnotes
\end{table}

@@

\begin{figure*}[t]
  \includegraphics[width=12cm]{figB1}
  \caption{Estimated begin (BGS) and end (EGS) of growing season for coniferous trees from MODIS Aqua/Terra net photosynthesis (PSN) product. A $1\times 1\,\unit{km}$ area around Svanhovd was selected. Daily averaged data for both 2018 and 2019 has been fitted with a quadratic polynomial function. The numerically computed root yields: BGS 122/106 day of year (\unit{doy}) and EGS 261/274 day of year for 2018/2019, respectively.}
  \label{fig:modis_Psn}
\end{figure*}
